### 2022-11-29 Draft Syllabus.Rmd
initial commit

---
title: "Degrees Draft Syllabus"
author: "Reggie Gilliard"
date: "11/29/2022"
output: html_document
---

# DegreesNYC Draft Syllabus {.tabset .tabset-fade .tabset-pills}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lessons {.tabset .tabset-fade .tabset-dropdown}

### Lesson 1 - An introduction to R (what, why, how)

1. What is R?
2. Why use R?
3. How do I install R?
4. The tools of the class.

```{r lesson1}

```

### Lesson 2 - Variables, data types, and operators

1. What are variables and why are they useful?
2. Assigning variables in R
3.

```{r lesson2}

```