### 2022-11-30 Draft Syllabus.Rmd
init: developing syllabus

---
title: "Degrees Draft Syllabus"
author: "Reggie Gilliard"
date: "11/29/2022"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Lessons {.tabset .tabset-fade .tabset-dropdown}

### Lesson 1 - An introduction to R (what, why, how) {.tabset .tabset-fade .tabset-pills}

#### Topics Covered

1. What is R?
2. Why use R?
3. How do I install R?
4. The tools of the class.

#### Lesson

##### What is R?

- Programming language used mostly for statisical analysis, data science, and data visualization
<!-- FOR REGGIE: List other common uses -->
- Many other uses

##### Why use R?

![R is a very common language in data science](./Lesson 1/r_prevalence.png)

#### Helpful Links

```{r lesson1}

```

### Lesson 2 - Variables, data types, and operators {.tabset .tabset-fade .tabset-pills}

#### Topics Covered

1. What are variables and why are they useful?
2. Assigning variables in R (=, <-, ->)
3. Data types
    + The six basic data types.
    + Character, numeric, and logical: the three most common types.
4. Mathematical operators
5. Logical operators

#### Helpful Links

1. [You can read more about the data types that we do not cover here.](https://www.programiz.com/r/data-types)

```{r lesson2}

```

### Lesson 3 - Tidy data and dplyr {.tabset .tabset-fade .tabset-pills}

#### Topics Covered

1. The dplyr package
2. Working in dplyr
    + The pipe (%>%)
    + Other useful dplyr functions
3. Tidy data
    + What is tidy data?
    + Why is it useful?

```{r lesson3}

```

### Lesson 4 - Vectors and lists {.tabset .tabset-fade .tabset-pills}

#### Topics Covered

1. What is a vector, what is a list?
    + What is the difference? 
<!-- (for Reggie: https://stackoverflow.com/questions/8594814/what-are-the-differences-between-vector-and-list-data-types-in-r) -->
2. Why are vectors useful?
3. Vectorized operations

```{r lesson4}

```

### Lesson 5 - Conditionals and loops {.tabset .tabset-fade .tabset-pills}

#### Topics Covered

1. If/then conditionals and their use
    + When to use them
2. While loops
    + When to use
3. For loops
    + When to use

```{r lesson5}

```

### Lesson 6 - Functions {.tabset .tabset-fade .tabset-pills}

1. What is a function?
2. When to use functions.
3. Writing functions in R.

```{r lesson6}

```