### 2023-05-13 Project Examples/Programs/project_examples.Rmd
Add non-shiny example

---
title: "Project Examples"
output: html_document
date: "2023-05-01"
---

# Setup
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
```

## Packages
```{r packages}
pkgs <- c("tidyverse", "kableExtra", "plotly")
invisible(lapply(pkgs, library, character.only = T))
```

## Paths
```{r paths}
root <- "D:/Documents/DegreesNYC/Degrees_data/Project Examples"
indir <- file.path(root, "Data")
intdir <- file.path(root, "Intermediate")
outdir <- file.path(root, "Output")
```

# Project

## Load Data

```{r load_data}
# data <- read.csv(file.path(indir, "mock_data.csv"))
data <- readxl::read_xlsx(file.path(intdir, "nyc_student_survey_2022.xlsx"), sheet = "Total")
data
```

## clean Data

```{r clean_data}
data %>%
  # Keep only Manhattan schools
  filter(grepl("\\d{2}M\\d{3}", DBN)) %>%
  select(-c(ends_with("Score"))) %>%
  # Extract district
  mutate(district = substr(DBN, 1, 2)) %>%
  {data <<- .}
```

## Descriptive Statistics

```{r descriptive_statistics}

dstats <- function(x, y) {
  name <- str_remove_all(y, "\\s")
  mean_name <- paste0(name, "_mean")
  min_name <- paste0(name, "_min")
  max_name <- paste0(name, "_max")
  x %>%
    group_by(district) %>%
    summarize(mean_name = mean(!!sym(y)),
              min_name = min(!!sym(y)),
              max_name = max(!!sym(y))) %>%
    setNames(c("District", mean_name, min_name, max_name))
}

map_dfc(c("Total Parent Response Rate", "Total Student Response Rate", "Total Teacher Response Rate"), function(x) dstats(data, x)) %>%
  select(-matches("District...[^1]")) %>%
  rename(District = `District...1`) %>%
  kable() %>%
  kable_paper()

data_dist <- map_dfc(c("Total Parent Response Rate", "Total Student Response Rate", "Total Teacher Response Rate"), function(x) dstats(data, x)) %>%
  select(-matches("District...[^1]")) %>%
  rename(District = `District...1`)
```

## Visualizations

```{r bargraph, fig.dim=c(10, 8)}

data_dist %>%
  rename_with(~ str_to_title(gsub("(Total)(\\w+)(Response)(Rate)_(\\w+)$", "\\1 \\2 \\3 \\4 \\5", .x, perl = T)), starts_with("Total")) %>%
  select(District, `Total Student Response Rate Mean`) %>%
  ggplot(aes(x = District, y = `Total Student Response Rate Mean`*100, fill = District, label = round(`Total Student Response Rate Mean`*100, 2))) +
  geom_bar(stat = "identity") +
  geom_text(vjust= -.5) +
  scale_y_continuous(limits = c(0, 100)) +
  ggtitle("Mean Student Response Rate by District") +
  ylab("Response Rate") +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank())

```

```{r heatmap, fig.dim=c(10, 8)}

heatmap <- data_dist %>%
  group_by(District) %>%
  rename_with(~ str_to_title(gsub("(Total)(\\w+)(Response)(Rate)_(\\w+)$", "\\1 \\2 \\3 \\4 \\5", .x, perl = T)), starts_with("Total")) %>%
  select(District, ends_with("Mean")) %>%
  pivot_longer(ends_with("Mean"), names_to = "Outcome", values_to = "Value") %>%
  ggplot(aes(x = District, y = Outcome, fill = Value)) +
  geom_tile() +
  ggtitle("Mean Response Rate by District")

ggplotly(heatmap)

```